### 2014-04-18 latex/dgci2014/rebuttal.txt
rebuttal

        First of all, we would like to thank the reviewers for their
comments. In the following, We will comment some of major remarks
(please note that all remaining remarks, typos and suggestions will be
considered in the camera ready version if the paper is finally
accepted).

**Path based vs. Euclidean Metrics and quadratic complexity**

The first objective of the paper is to propose an algorithmic solution
to efficiently compute distance transformation for some path based
norms which mimics the parallel and separable approach used for
Euclidean and $L_p$ metrics. We will double-check the introduction to
make it clear that the paper focuses on that point.

Concerning the discussion about path based metric vs. Euclidean one,
we are convinced that path based norms are very interesting
mathematical objects and crucial in some applications (non-convex
domains, streaming DT, when the anisotropic nature of their unit ball
is consistent with the application, e.g. in crystallography...).

However, the huge literature about path-based metric optimization to
minimize the rotational dependency tends to illustrate that the
approximation quality (w.r.t. Euclidean distance) also matters. In this
context, the size of the mask is an important parameter (together with
optimized weights). Introduction will be rephrased to clarify this
context.

It is out of the scope of the paper (and won't be added in the
introduction) but please note that for medial axis extraction purposes
(defined as centers of maximal balls), linear  time algorithms
exists for the Euclidean metric [Coeurjolly-PAMI 2007, Hesselink PAMI
2008..]. For path based norms, pre-computations and look-up tables are
usually considered. As far as we know, it is still an active open
problem to prove that distances on shortest paths can be used to
extract the set of (or a subset of) maximal ball centers for the
metric [Strand, DGCI 2011].

**Experiments**

 - Graph size: we agree that graphs labels are quite small. We will
   correct this for the final version if the paper is accepted.

 - Scale-up w.r.t. to number of threads:  definitely
   doable. Experiments will be done and could be included in the final
   version (depending on the final version page limit).

 - Timings for low-size mask: zoom-in to the left-part of the graph
   can be added.

 - Comparison w.r.t. double-scan approach: doable too. However, we may
   end-up with issues with respect to page limit.

**Spelling, Grammar, Paper structure**

All apologizes for the bad English. The text will be double-checked
for the final version.